The Creation of a Programming Language

### 2022-10-29 paper.tex
continue writing paper

% hyperref abbreviations
\newcommand{\hr}[2]{\hyperref[#2]{#1}}
\newcommand{\hrc}[1]{\hyperref[#1]{#1}}

@@

% Include C code from file #1 from lines #2 to #3
\newcommand{\sourcecode}[3] {
    \ver{
    \lstset{basicstyle=\scriptsize,keywordstyle=\color{blue},stepnumber=5,
        stringstyle=\color[rgb]{.3,.5,.1},commentstyle=\color{gray},
        morecomment=[l][\color{magenta}]{\#},
        postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space}}
    \lstinputlisting[language=C,numbers=left,breaklines=true,
        frame=single,title=#1,firstnumber=#2,firstline=#2,lastline=#3]{#1}
    }
}

@@

    \begin{minipage}{\linewidth}

    \lstset{basicstyle=\scriptsize,keywordstyle=\color{blue},stepnumber=5,
        stringstyle=\color[rgb]{.3,.5,.1},commentstyle=\color{gray},
        morecomment=[l][\color{magenta}]{\#},
        postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space}}
    \lstinputlisting[language=C,numbers=left,breaklines=true,
        frame=single,title=\hrc{#1},
        firstnumber=#2,firstline=#2,lastline=#3]{#1}

@@

\tableofcontents\newpage

@@

\section{Implementation}
\subsection{Used Software}
At the heart of making any project in the C programming language
stands the C compiler. This project uses
\emph{gcc}, the GNU Compiler Collection, which, as the name suggests, also 
supports more languages than solely C. Substantiating the development process is 
\emph{GNU Make}, a general-purpose build system. It handles tasks like building
the final executable program step-by-step from the project source code and directory clean-up.
The C ecosystem is very easy to use on and fundamentally supported by 
\emph{GNU+Linux}, the operating system used to program on.
The code editor used was \emph{Microsoft Visual Studio Code}, providing great
help with built-in git integration and auto-completion features.
The tools \emph{Valgrind} and \emph{gdb}, the GNU Project Debugger were of great help
when trying to find of errors and memory leaks, see section \ref{memleaks} for details.
\emph{Git} is the project management and version tracking software used to back up and share
the project files on \emph{GitHub.com}, a Microsoft hosted git repository and
web front-end to store and collaborate on source code.

@@

Here is a file tree overview of the project which will be referenced
several times in the following sections of this paper. In the depicted file tree
only source files are shown. No build files, executable files or files that
are of no further importance regarding \name are depicted.
\ver{\begin{flushleft}
    \scriptsize\setlength{\DTbaselineskip}{7pt}
    \dirtree{%
        .1 project root.
        .2 \hr{.gitignore}{.gitignore}.
        .2 \hr{README.md}{README.md}.
        .2 \hr{Makefile}{Makefile}.
        .2 \hr{mapop.gperf}{mapop.gperf}.
        .2 pseudocode.
            .3 \hr{parser.c}{pseudocode/parser.c}.
        .2 examples.
            .3 \hr{mandelbrot.nts}{examples/mandelbrot.nts}.
            .3 \hr{fibonacci.nts}{examples/fibonacci.nts}.
            .3 \hr{exec.nts}{examples/exec.nts}.
        .2 src.
            .3 \hr{main.c}{src/main.c}.
            .3 interpreter.
                .4 \hr{bytecode.c}{src/interpreter/bytecode.c}.
                .4 \hr{bytecode.h}{src/interpreter/bytecode.h}.
                .4 error.
                    .5 \hr{error.c}{src/interpreter/error/error.c}.
                    .5 \hr{error.h}{src/interpreter/error/error.h}.
                .4 \hr{interpreter.c}{src/interpreter/interpreter.c}.
                .4 \hr{interpreter.h}{src/interpreter/interpreter.h}.
                .4 libraries.
                    .5 \hr{libraries.c}{src/interpreter/libraries/libraries.c}.
                    .5 \hr{libraries.h}{src/interpreter/libraries/libraries.h}.
                    .5 stdlib.
                        .6 \hr{stdlib.c}{src/interpreter/libraries/stdlib/stdlib.c}.
                        .6 \hr{stdlib.h}{src/interpreter/libraries/stdlib/stdlib.h}.
                .4 \hr{localizer.c}{src/interpreter/localizer.c}.
                .4 \hr{localizer.h}{src/interpreter/localizer.h}.
                .4 mappings.
                    .5 \hr{operations.c}{src/interpreter/mappings/operations.c}.
                    .5 \hr{operations.h}{src/interpreter/mappings/operations.h}.
                    .5 \hr{mapop.c}{src/interpreter/mappings/mapop.c}.
                    .5 \hr{mapop.h}{src/interpreter/mappings/mapop.h}.
                .4 memory.
                    .5 \hr{hashtable.c}{src/interpreter/memory/hashtable.c}.
                    .5 \hr{hashtable.h}{src/interpreter/memory/hashtable.h}.
                    .5 \hr{array.c}{src/interpreter/memory/array.c}.
                    .5 \hr{array.h}{src/interpreter/memory/array.h}.
                    .5 \hr{stack.c}{src/interpreter/memory/stack.c}.
                    .5 \hr{stack.h}{src/interpreter/memory/stack.h}.
                .4 \hr{parser.c}{src/interpreter/parser.c}.
                .4 \hr{parser.h}{src/interpreter/parser.h}.
                .4 processing.
                    .5 \hr{implementations.c}{src/interpreter/processing/implementations.c}.
                    .5 \hr{implementations.h}{src/interpreter/processing/implementations.h}.
                .4 \hr{runtime.c}{src/interpreter/runtime.c}.
                .4 \hr{runtime.h}{src/interpreter/runtime.h}.
                .4 \hr{tokenizer.c}{src/interpreter/tokenizer.c}.
                .4 \hr{tokenizer.h}{src/interpreter/tokenizer.h}.
    }
\end{flushleft}}

@@

\subsection{Implementing \name{}}
This section will explain the process of constructing and the inner workings of
the \name{} interpreter. It is therefore the most complicated and detailed section
in this paper. Examples will be made with both pseudo code and actual source
code from the project to further illustrate the decisions made during
interpreter design. The file paths used are relative to the project's root directory.

@@

Following is a very quick summary of the few 
things happening in the \hrc{src/main.c} file.
First, the \hrc{src/interpreter/interpreter.h} header file is included. 
The main function then decides whether to
read a program from a file or from standard input. 
At last, the function \emph{interpret}
from the said included header is called with the selected input.

@@

\subsubsection{Lack of a Development Environment}
This is probably the greatest problem of \name{} and any new prorgramming
languages in general.

@@

\subsection{List of Figures}\listoffigures
\vspace{1cm}
Figure \ref{f_mandelbrot} was generated using a \name{} program, the source code
can be found in the file \hrc{examples/mandelbrot.nts}.
Figure \ref{f_diagram} was created using an online diagram creation tool available
on \url{https://app.diagrams.net}.
[\cte{drawio}]

\section{Declaration of Authenticity}
Hereby I, Florian Donnelly, declare to have written this paper independently, 
also including the entire source code of \name{} by myself using resources 
listed under the \emph{Bibliography} section on page \pageref{bibliography}.
% TODO: check correctness of all file paths!! e.g. src/...

>>>>

\definecolor{mygreen}{rgb}{0.3, 0.5, 0.2}
\definecolor{myteal}{rgb}{0.1, 0.2, 0.3}
\definecolor{myblue}{rgb}{0.2, 0.0, 0.6}
\definecolor{mypink}{rgb}{0.8, 0.3, 0.6}
\definecolor{myorange}{rgb}{0.7, 0.4, 0.1}

% Use fancier hyperrefs
\hypersetup{
    colorlinks=true,
    pdfborder={3 3 .2},
    citecolor=mygreen,
    linkcolor=myteal,
    urlcolor=myblue,
}

% hyperref abbreviations
\makeatletter
\DeclareRobustCommand{\filename}[1]{%
    \begingroup
    \def\textendash{-}%
    \filename@parse{#1}%
    \edef\filename@base{\detokenize\expandafter{\filename@base}}%
    \texttt{\filename@base.\filename@ext}%
    \endgroup
}
\makeatother
\newcommand{\hr}[2]{\hyperref[#2]{#1}}
\newcommand{\hrc}[1]{\hyperref[#1]{\filename{#1}}}

@@

% Include C code from file #1 from lines #2 to #3
\lstset{basicstyle=\scriptsize,keywordstyle=\color{myorange},stepnumber=5,
        frame=none,keepspaces=false,
        rulecolor=\color{black},tabsize=2,
        stringstyle=\color{mygreen},commentstyle=\color{mypink},
        morekeywords={NULL},
        morecomment=[l][\color{magenta}]{\#},
        postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space}}

\newcommand{\sourcecode}[3] {
    \ver{
        \lstinputlisting[language=C,numbers=left,breaklines=true,
        frame=single,title=#1,firstnumber=#2,firstline=#2,lastline=#3]{#1}
    }
}

@@

    \begin{minipage}{\linewidth}
    \lstinputlisting[language=C,numbers=left,breaklines=true,
        frame=single,title=\hr{#1}{#1},
        firstnumber=#2,firstline=#2,lastline=#3]{#1}

@@

{
    \hypersetup{hidelinks}
    \tableofcontents\newpage
}

@@

\section{Implementation}
\subsection{Used Software}
At the heart of making any project in the C programming language
stands the C compiler. This project uses
\emph{gcc}, the GNU Compiler Collection, which, as the name suggests, also 
supports more languages than solely C. Substantiating the development process is 
\emph{GNU Make}, a general-purpose build system. It handles tasks like building
the final executable program step-by-step from the project source code and directory clean-up.
The C ecosystem is very easy to use on and fundamentally supported by 
\emph{GNU+Linux}, the operating system used to program on.
The code editor used was \emph{Microsoft Visual Studio Code}.
It has built-in git integration and auto-completion features.
The tools \emph{Valgrind} and \emph{gdb}, the GNU Project Debugger, were of help
when trying to find of errors and memory leaks, see section \ref{memleaks} for details.
\emph{Git} is the project management and version tracking software used to back up and share
the project files on \emph{GitHub.com}, a Microsoft hosted git repository and
web front-end to store and collaborate on source code.

@@

Here is a file tree overview of the project which will be referenced
several times in the following sections of this paper. In the depicted file tree
only source files are shown. No build files, executable files or files that
are of no further importance regarding \name are depicted.
\ver{\begin{flushleft}
    \scriptsize\setlength{\DTbaselineskip}{7pt}
    \dirtree{%
        .1 project root.
        .2 \hrc{.gitignore}.
        .2 \hrc{README.md}.
        .2 \hrc{Makefile}.
        .2 \hrc{mapop.gperf}.
        .2 pseudocode.
            .3 \hrc{pseudocode/parser.c}.
        .2 examples.
            .3 \hrc{examples/mandelbrot.nts}.
            .3 \hrc{examples/fibonacci.nts}.
            .3 \hrc{examples/exec.nts}.
        .2 src.
            .3 \hrc{src/main.c}.
            .3 interpreter.
                .4 \hrc{src/interpreter/bytecode.c}.
                .4 \hrc{src/interpreter/bytecode.h}.
                .4 error.
                    .5 \hrc{src/interpreter/error/error.c}.
                    .5 \hrc{src/interpreter/error/error.h}.
                .4 \hrc{src/interpreter/interpreter.c}.
                .4 \hrc{src/interpreter/interpreter.h}.
                .4 libraries.
                    .5 \hrc{src/interpreter/libraries/libraries.c}.
                    .5 \hrc{src/interpreter/libraries/libraries.h}.
                    .5 stdlib.
                        .6 \hrc{src/interpreter/libraries/stdlib/stdlib.c}.
                        .6 \hrc{src/interpreter/libraries/stdlib/stdlib.h}.
                .4 \hrc{src/interpreter/localizer.c}.
                .4 \hrc{src/interpreter/localizer.h}.
                .4 mappings.
                    .5 \hrc{src/interpreter/mappings/operations.c}.
                    .5 \hrc{src/interpreter/mappings/operations.h}.
                    .5 \hrc{src/interpreter/mappings/mapop.c}.
                    .5 \hrc{src/interpreter/mappings/mapop.h}.
                .4 memory.
                    .5 \hrc{src/interpreter/memory/hashtable.c}.
                    .5 \hrc{src/interpreter/memory/hashtable.h}.
                    .5 \hrc{src/interpreter/memory/array.c}.
                    .5 \hrc{src/interpreter/memory/array.h}.
                    .5 \hrc{src/interpreter/memory/stack.c}.
                    .5 \hrc{src/interpreter/memory/stack.h}.
                .4 \hrc{src/interpreter/parser.c}.
                .4 \hrc{src/interpreter/parser.h}.
                .4 processing.
                    .5 \hrc{src/interpreter/processing/implementations.c}.
                    .5 \hrc{src/interpreter/processing/implementations.h}.
                .4 \hrc{src/interpreter/runtime.c}.
                .4 \hrc{src/interpreter/runtime.h}.
                .4 \hrc{src/interpreter/tokenizer.c}.
                .4 \hrc{src/interpreter/tokenizer.h}.
    }
\end{flushleft}}

@@

\subsection{Implementing \name{}}
This section will explain the process of constructing and the inner workings of
the \name{} interpreter. It is therefore the most complicated and detailed section
in this paper. Examples will be made with both pseudo code and actual source
code from the project to show the decisions made during
interpreter design. Only file names are used, as those are unique. To view the
entire file path, look up the file names in the file tree on page \pageref{FileStructure}.

@@

Following is a very quick summary of the few 
things happening in the \hrc{src/main.c} file.
First, the \hrc{src/interpreter/interpreter.h} header file is included. 
The main function then decides whether to
read a program from a file or from standard input. 
At last, the function \emph{interpret}
from the included header is called with the selected input.

@@

\subsubsection{Lack of a Development Environment}
This is probably the largest problem of \name{} and new programming
languages in general.

@@

\subsection{List of Figures}
{
    \hypersetup{hidelinks}
    \listoffigures
}
\vspace{1cm}
Figure \ref{f_mandelbrot} was generated using a \name{} program, the source code
can be found in the file \hrc{examples/mandelbrot.nts}.
Figure \ref{f_diagram} was created using an online diagram creation tool available
on \url{https://app.diagrams.net}.
[\cte{drawio}]

\section{Declaration of Authenticity}
Hereby I, Florian Donnelly, declare to have written this paper independently, 
also including the entire source code of \name{}, using resources 
listed under the \emph{Bibliography} section on page \pageref{bibliography}.
% TODO: check correctness of all file paths!! e.g. src/...
